---
output:
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: ../../svm-latex-ms.tex
#  word_document:
#always_allow_html: yes
title: "Regime Change and Information Control"
thanks: " "
author:
- name: Ross Dahlke
affiliation: Stanford University
abstract: ""
keywords: ""
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
spacing: double
bibliography: bibliography.bib
biblio-style: apa
header-includes: \usepackage{graphicx}
---

```{r echo = F, message = F}
knitr::opts_chunk$set(include = F, echo = F, warning = F, message = F)
```

```{r include = T}
# wordcountaddin:::text_stats()
```

Among comparative political scientists, the study of regimes around the world centers on the question: What leads to countries adopting democracy? In comparative politics literature, attempts at answering this question primarily start by examining the relationship between economic development and democratization. I start by examining the economic argument and then move onto non-economic explanations of regime change.

This paper summarizes literature on a core subject of comparative politics: regimes and regime change. In addition, I add a perspective on how the literature relates to censorship, propaganda, and information control. In total, the main goal of this final paper is to engage with a fundamental topic in comparative politics and layer on top some ideas from the field of mass communication.

# Economic Development and Democracy

There is a strong correlation between levels of economic development and democracy in a country. However, there has been a historical divide on the causality of economic development and the creation or maintenance of democracy. Debates on the importance of economics in democratization were distilled into two arguments by Przeworski and Limongi [-@przeworski], the "endogenous" or "exogenous" models. In the endogenous view, economic development spurs the emergence of democracy. In the exogenous model, democracy emerges independent of economic development but is more likely to survive in economically developed countries. In other words, economic development either _causes_ democratization or _maintains_ democratization caused by other processes.

Modernization is the focal point of the endogenous view of economic development and democracy. Modernization of countries drives a sequence of events of "industrialization, urbanization, education, communication, mobilization, and political incorporation" [@przeworski]. Under an American- and democracy-centric view of global development, modernization is an accumulation of progressive social changes that readies a nation for its final culmination: democracy. Under the endogenous theory of democratization, there is a threshold of economic development at which we would expect authoritarian regimes to transition to democracy. 

Sociologists point out how modernization develops communicative infrastructure. This infrastructure can facilitate democratization. For example, Castells [-@castells] refers to the output of modernization as the rise of network society. As the world globalizes, it is dominated by central nodes which act as a connector for multiple types of networks. For example, New York City is a central hub for financial, news, and entertainment networks. In order to modernize, a country must accept its position within the networked world. Inevitably, this network position opens up the country to more information flows which leads to democratization. 

In contrast, the exogenous view of democratization posits economic development not cause of democracy but instead as a maintainer of democracy. Lipset [-@lipset] argues that the emergence of democracy is independent of economic factors. However, the likelihood of a regime surviving increases in more developed nations. Instead of the threshold of democratization offered by the endogenous view, there is a threshold of stability for regimes. There is a "U" shape that develops on the likelihood of a dictatorship transitioning to democracy where authoritarian regimes with low and high levels of modernization are likely to persist. For example, modern economic development in China has not democratized the nation and has instead cemented the Chinese Communist Party's rule over the country. This exogenous view has withstood empirical scrutiny whereas the endogenous view has not. 

Put together, there is little causal explanatory power of economic development bringing down authoritarian regimes and transitioning them to democracy. Recent examples such as China, Singapore and Rwanda have experienced dramatic recent economic growth but remain authoritarian states at various levels. Instead, once established, democracies that are wealthier and more modernized are more likely to survive. Further, Lipset theorized that democracies are more likely to be destabilized when their economies grow quickly. These accelerative periods breed "extremist moments." For example, fascism and communism were products of rapid economic development. However, the evidence suggests that instead of authoritarianism growing during economic advancement that democracies fall during economic contractions [@przeworski]. 

In summary, economic development is crucial towards maintaining democracy but does not itself cause democratic transition. There are other factors that cause regime change. Instead, the level of modernization predicts whether the new government will be maintained. Therefore, researchers moved to examine predictors of democratization other than economic growth.

# Other Factors of Regime Change

If economic development does not cause regime change, what does? Robinson and Acemoglu [-@robinson] place the struggle between elites and citizens as central to the development of democracy. Rustow [-@rustow] summarizes a number of explanations for democratization including features of the conflict and reconciliation of groups in a society that form social and political structure and the need for certain beliefs and psychological attitudes among the citizenry. McFaul analyzes democratization from the historical perspective of the third wave of democratization and the inherent openness to democracy in certain cultures. Huntington offers the perspective of the fourth wave of democratization in a postcommunist world and the significance of elites' willingness to democratize a nation.

# Balance of power between elites and citizens

While it is not economic development _per se_ that causes nondemocracies to transition to democracy, one possibility is social class struggle between elites and citizens drives democratization [@robinson]. In this view, these two groups have well-defined and divergent economic preferences. Elites in a country want to maintain authoritarian rule because it provides them with higher levels of economic wealth. Whereas the majority of citizens want democratic institutions under the assumption that they will gain economic prosperity from a government restructure. The dueling preferences of these two groups are resolved through an inherently conflictual political process where de jure and de facto political power are wielded against one another. The balance of power between elites and citizens determines whether a society transitions from nondemocracy to democracy and perhaps if it reverts back to nondemocracy later. 

The factors that Robinson [-@robinson] discusses that can alter the levels of political power controlled by citizens and elites can be categorized into three levels: individual-level, country-level, and macro-level. I also offer information flow and communication as a meta-level. The balance of power is determined by these factors that dictate whether there will be regime change in a country.

## Individual-level factors

First, there are individual-level factors that alter the balance of power in a country. Civil society and the effective threat of revolution, particularly the level of self-organization of citizens, can shift the balance of power towards democracy. Relatedly, political identities can create citizens with varying levels of solidarity. Or, a large middle class can act as a buffer of conflict between elites and the citizens. 

## Country-level factors

Next, country-level factors help to determine the balance of power in a country. The sources of income and composition of wealth, such as elites being heavily invested in land versus physical and human capital, can create power imbalances that make democratization more likely. Political institutions can preemptively use repression to avoid democratization. Another country-level factor is inter-group inequality. A citizenry divided along racial or religious lines can make democratization more-or-less likely.

## Macro-level factors

Finally, there are macro-level events and trends that shift the balance of power between citizens and elites. Exogenous shocks and crises in a country can create a transition of de facto power to citizens and make revolution less costly to carry out. Globalization can also shift elite exposure to international financial markets and fundamentally shape economies and economic power. 

## Meta-level factors

From a communication studies perspective, information flows can also act as a sort of meta-level in the balance of power between citizens and elites. Free-flowing information can help citizens organize politically to increase their de facto power. Authoritarian regimes participate in various types of censorship in order to quell citizens' de facto power, such as China's censorship of online content that produces collective action [@king]. Regimes also can participate in propaganda in order to increase the legitimacy of their government in order to communicate its de jure power. A regime can participate in propaganda as persuasion in order to maintain devotion to and legitimacy of the nation, for example, North Korea's anti-American and pro-state propaganda. Or, a regime can use propaganda as signaling simply to show that it has the power to maintain compulsory exhibitions of pro-regime attitudes, such as in the case of Syria and public displays of loyalty.

# Conditions for Democracy

Rustow [-@rustow] offers conditions under which democracy can thrive. One argument centers around politicians and the other around the citizenry. These arguments made by Rustow relate to historical analyses by McFaul's analysis on the fourth wave of democracy and focuses on elites and Huntington's research on the third wave of democracy, which he centers on individualistic features.

## Politicians

The politician-centered explanation offered by Rustow is similar to Robinson's [-@robinson] in that social conflict and reconciliation are essential components to democracy. For example, there is a vitality of American institutions because of citizens' "multiple memberships in potential groups" which spurs politicians to also join these groups. These crosscutting associations mediate American politics and allow different coalitions to emerge within the democratic process--a process that is self-reinforcing. In addition, Rustow points to arguments that democratic stability requires a commitment to democratic values among professional politicians. These democratic values are created and maintained through politicians' links to one another through ties of political organizations. 

McFaul's [-@mcfaul] analysis of the fourth wave of democracy (1990-present), or democratization in a postcommunist world, also hinges on elites and the tenet of the balance of power dictating transitions to and from democracy. Decommunization triggered this fourth wave of regime change, but unlike prior waves, the fourth wave produced regime change both to democracy _and_ dictatorship. The ideological orientation of the more powerful party that enjoyed an asymmetrical balance of power largely determined the type of regime that emerged out of the fourth wave. Therefore, democracy did emerge in some countries, but it was only where pro-democracy forces enjoyed a decisive power advantage. 

## Citizens

Another potential explanation put forth by Rustow is that there is a need for certain beliefs or psychological attitudes among citizens. In other words, there is a certain culture of democracy that is needed among the masses in order to maintain a democracy. There either needs to be a consensus in common beliefs and values or in the procedure to the rules of the game. There is an "agreement to differ." There are potentially deeper psychological and civic attitudes such as empathy and a willingness to participate that underline this culture of democracy. However, it is not just participation, but also "traditional or parochial" attitudes that are the bedrock of a "civic culture" that is necessary for democracy. 

Propaganda can help to shape citizens' psychological attitudes in relation to the values necessary for democracy. A regime can participate in propaganda to create or reinforce nondemocratic attitudes. In addition, it can censor information that argues against the regime. For example, Saudi Arabia silencing influential anti-regime activists [@pan2020].

Huntington's [-@huntington] analysis of "Democracy's Third Wave" (1974-1990) also points to culture among citizens as being one of the obstacles to democratization. Cultural traditions set the attitudes, values, beliefs, and related behavior patterns that determine whether a society is conducive to democracy or is antidemocratic. Specifically, Huntington points to Confucianism and Islam as two cultures which are inherently antidemocratic. He offers that Confucianism is a culture that emphasizes the collective over the individual, and there is a lack of traditional rights against the state. Unlike America, which has a notion of "God-given" rights which are unalienable, Confucianism's rights are created by the state. Huntington also points to Islam as an inherently antidemocratic culture because of the lack of separation between church and state. In Islam, the leader of the church is also the leader of the state. Therefore, it is only a divine power, not the consent of the governed, that allows for a leader. 

Both of these examples are inductive. It is easy to look back on history and create strings that tell a story of the present. The same exercise could be applied to post-World War II Germany and conclude that German culture is not conducive to democracy. Comparative political scientists pointed to historical disparities in regional development and culture-based class alliances as shaping a political culture that was not aligned with the possibility of democracy, just like Huntington concludes that cultures like Confucuianism and Islam are not applicable to democratic societies [@bernhard]. In the present, the argument that Germany cannot sustain democracy because of its culture seems unfounded. By the 1980s a new generation of Germans shared values that converged with other western democracies and remains a strong European democracy [@bernhard].

# Conclusion

Historically, the debate around democratization in comparative politics has centered around economic development. Economic development may not cause democratization, but development can maintain democracies that are established for other reasons. Some of the reasons that democracy can be established are the power balance between citizens and elites, psychological attitudes and beliefs towards democracy, and other cultural factors. Further, the flow of information, including information control via propaganda and censorship, plays a crucial role in these democratizing factors.

# References