---
output:
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: ../../svm-latex-ms.tex
#  word_document:
#always_allow_html: yes
title: "Regime Change and Information Control"
thanks: " "
author:
- name: Ross Dahlke
affiliation: Stanford University
abstract: "This paper summarizes literature on a core subject of comparative politics: regimes and regime change. In addition, I add a perspective on how the literature relates to censorship, propaganda, and information control. In total, the main goal of this final paper is to give me an opportunity to learn more about the fundamentals of comparative politics and layer on top some ideas from the field of mass communication."
keywords: ""
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
spacing: double
bibliography: bibliography.bib
biblio-style: apsr
header-includes: \usepackage{graphicx}
---

```{r echo = F, message = F}
knitr::opts_chunk$set(include = F, echo = F, warning = F, message = F)
```

```{r include = T}
wordcountaddin:::text_stats()
```

Democracy is a founding principle in the United States and has persisted through its history. This collective value has not only meant that American have taken their own democracy for granted, it also leads to an implicit worldview that centers democracy as the ultimate goal for countries around the world. As a result, political scientists have historically framed the study of regimes around the world around the question: What leads to countries adopting democracy? [Need coherent thesis]

# Economic Development and Democracy

There is a strong correlation between levels of economic development and democracy in a country. However, there has been a historical divide on the causality of economic development and the creation or maintenance of democracy. Przeworski and Limogni [-@przeworski] distill two arguments, either economic development is "endogenous" and spurs the emergence of democracy; or it is "exogenous," and democracy emerges independent of economic development but is more likely to survive in economically developed countries. In other words, economic development either causes democratization or maintains democratization caused by other processes.

The endogenous view of economic development and democracy is one of modernization. Modernization of countries drives a sequences of events of "industrialization, urbanization, education, communication, mobilization, and political incorporation" [@przeworski]. Ultimately, modernization is an accumulation of progressive social changes that readies a nation for it's final culmination: democracy. Under the endogenous theory of democratization, there is a threshold of economic development at which we would expect authoritarian regimes to transition to democracy. However, this theory does not hold up under empirical testing. Instead, evidence points towards the exogenous view of economic development and democracy.

The exogenous view of democratization posits economic development not as the driver towards democracy but instead as a maintainer of democracy. Lipset [-@lipset] argues that the emergence of democracy is independent of economic factors. However, the likelihood of a regime surviving increases in more developed nations. Instead of the threshold of democratization offered by the endogenous view, there is a threshold of stability for regimes. There is a "U" shape that develops on the likelihood of a dictatorship transitioning to democracy where authoritarian regimes with low and high levels modernization are likely to persist. 

Put together, there is little causal explanatory power of economic development bringing down authoritarian regimes and transitioning them to democracy. Instead, once established, democracies that are wealthier and more modernized are more likely to survive. Further, Lipset even went so far as to theorize that democracies are more likely to be destabilized when their economies grow quickly. These accelerative periods breed "extremist moments." For example, fascism and communism were produts of rapid economic development. However, it appears more likely that instead of authoritarianism growing during economic advancement that democracies fall during economic contractions [@przeworski]. 

In summary, economic development does not breed democracies. In contrast, there are other factors that cause regime change. It is the level of modernization that predicts whether the new government will be maintained. In addition, economic crises can lead to democracies falling into authoritarian states. 

# Factors of Regime Change

If economic development does not cause regime change, what does? Robinson and Acemoglu [-@robinson] place the struggle between elites and citizens as central to the development of democracy. Rustow [-@rustow] summarizes a number of explanations for democratization including features of the conflict and reconciliation of groups in a society that form social and political structure and the need for certain beliefs and psychological attitudes among the citizenry. 

While it is not economic development _per say_ that causes nondemocracies to transition to democracy, one possibility is that social struggle, largely economic, between elites and citizens drives democratization [@robinson]. In this view, these two groups have well-defined economic preferences. Elites in a country want to maintain authoritarian rule because it provides them with higher levels of economic wealth. Whereas the majority of citizens want democratic institutions under the assumption that they benefit from them and therefore will exert their power in an effort to attain democracy. The dueling preferences of these two groups are resolved through an inherently conflictual political process where de jure and de facto political power is wielded against one another. The balance of power between elites and citizens determines whether a society transitions from nondemocracy to democracy and perhaps if it reverts back to nondemocracy later. 

There are factors that can alter the levels of political power controlled by citizens and elites [@robinson]. Civil society and the effective threat of revolution, particularly the level of self-organization of citizens, can shift the balance of power towards democracy. Exogenous shocks and crises in a country can create a transition of de facto power to citizens and make revolution less costly to carry out. The sources of income and composition of wealth, such as elites being heavily invested in land versus physical and human capital, can create power imbalances that make democratization more likely. Political institutions can preemptively use repression to avoid democratization. Inter-group inequality, for example if a citizenry is divided along racial or religious lines, can make democratization more-or-less likely. A large middle class can act as a buffer of conflict between elites and the citizens. Globalization can shift elite exposure to international financial markets. And political identities can create citizens with varying levels of solidarity. All of these factors impact the balance of de facto and de jure power that can be used by citizens or elites. This balance of power, according to Robinson and Acemoglu [-@robinson], is what ultimately create transitions to democracy.

Rustow [-@rustow] offers potential explanations under which democracy can thrive. One of these explanations is similar to the democratization thesis from Robinsin [-@robinson] in that it insists that conflict and reconciliation of conflicts are essential components to democracy. For example, there is a vitality of American institutions because of citizens' "multiple memberships in potential groups." These crosscutting associations mediate American politics and allow different coalitions to emerge within the democratic process--a process that is self-reinforcing. In addition, Rustow points to arguments that democratic stability requires a commitment to democratic values among professional politicians. These democratic values are created and maintained through politicians' links to one another through ties of political organizations. 

McFaul's [-@mcfaul] analysis of the fourth wave of democracy, or democratization in a postcommunist world, also hinges on tenant of the balance of power dictating transitions to and from democracy. Decommunization triggered this fourth wave of regime change, but unlike prior waves, the fourth wave produce regime change both to democracy _and_ dictatorship. The ideological orientation of the more powerful party that enjoyed an asymmetrical balance of power in a country largely determined the type of regime that emerged out of the fourth wave. Therefore, democracy did emerge in some countries, but it was only where pro-democracy forces enjoyed a decisive power advantage. 

Another potential explanation put forth by Rustow is that there is a need for certain beliefs of psychological attitudes among citizens. In other words, there is a certain culture of democracy that is needed among the masses in order to maintain a democracy. There either needs to be a consensus in common beliefs and values or in the procedure to the rules of the game. There is an "agreement to differ." There are potentially deeper psychological and civic attitudes such as empathy and a willingness to participate that underline this culture of democracy. However, it just not participate, but also "traditional or parochial" attitudes that are the bedrock of a "civic culture" that is necessary for democracy.

Huntington's [-@huntington] analysis of Democracy's Third Wave also points to culture as being one of the obstacles to democratization. Cultural traditions set the attitudes, values, beliefs, and related behavior patterns that determine whether a society is conducive to democracy or is antidemocratic. Specifically, Huntington points to Confucianism and Islam as two cultures which are inherently antidemocratic. He says that Confucianism is a culture that emphasizes the collective over the individual and there is a lack of traditional rights against the state. Unlike America, which has a notion of "God-given" rights which are unalienable, Confucianism's rights are created by the state. Huntington also points to Islam as an inherently antidemocratic culture because of the lack of separation between church and state. In Islam, the leader of the church is also the leader of the state. Therefore, it is only a divine power, not the consent of the governed, that allows for a leader. Both of these examples feel to me to be too inductive. It is easy to look back on history and create strings that tell a story of the present. The same exercise could be applied to post-World War II Germany and conclude that German culture is not conducive to democracy. In the present, that argument that Germany cannot sustain democracy because of its culture seems unfounded.  

# References