---
output:
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: ../svm-latex-ms.tex
#  word_document:
#always_allow_html: yes
title: "Retail investors and media psychology"
thanks: " "
author:
- name: Ross Dahlke
affiliation: Stanford University
abstract: ""
keywords: ""
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
spacing: double
bibliography: bibliography.bib
biblio-style: apsr
header-includes: \usepackage{graphicx}
---

```{r echo = F, message = F}
knitr::opts_chunk$set(include = F, echo = F, warning = F, message = F)
```

```{r include = T}
wordcountaddin:::text_stats()
```

# Abstract

# Introduction

Financial markets were recently upended by massive short squeezes of a cadre of stocks including GameStop, AMC Entertainment Holdings, BlackBerry, and Nokia. The recent events saw historic fluctuations in the prices of these stocks. Individual investors made tens-of-millions of dollars, often with leveraged positions. What is particularly unique about these events is that they were triggered by retail investors. These retail investors, bolstered by financial stimuli and an abundance of free time due to the COVID-19 pandemic, organized collective behavior actions on an internet forum, the Reddit subforum r/WallStreetBets. 

Media commentators quickly offered various perspectives on the events. Is this class warfare? A populist movement? Just some bored, lonely young men trying to make money? There is a deep pool of academic literature that explores the psychological motivations of retail investors and personal finance more broadly. For example, _________. In addition, there is a small, yet robust and powerful set of research on the impacts of media on financial markets. However, I am interest in the media psychology component of financial decisions. I am interested in how media impact financial markets in-so-much-as how the media effects, and is driven by, individual investors. 

More broadly, my research investigates how individuals enter and participate in complex systems. In this case, the totality of the financial markets are beyond comprehension. Investors, particularly small, retail traders, are unlikely to ever fully understand the complex system of global finance. Yet, they do. And as is increasingly common, participants are coordinating activity in order to alter the systems that they participate in. As it pertains to retail investors, I ultimately want to know both _why_ they participate in the markets, and more importantly, _how_ they participate in the financial markets. What is the media ecology that individual investors are exposed to or seek out? How do novice traders come to trust the financial investment advice of various media outlets, communities, or individuals? What do people do after they lose their life savings in GameStop stock? How much time do retail investors devote to trading financial securities? 

## COVID-19 and Retail Traders

The COVID-19 global pandemic has spurred a rise in retail investors that rivals the boom of retail investors during the internet bubble after the rise of discount stock brokerages. At the center of the rise in retail investors is Robinhood. Robinhood is a smart phone app and stock trading platform that offers free trading of stocks and other securities. While other trading platforms have since adopted free trading, Robinhood created and maintained a significant following by being the first to offer free trading, it's gamified user interface, and a target demographic of young, new investors. With Robinhood easing barriers of entry for new investors, financial stimuli from the government due to the pandemic, and newly-found swathes of free-time, these retail investors grew in size and influence during the pandemic. 

Retail investors now routinely account for 20% of stock market activity [@eaton2021] These retail investors on Robinhood significantly impact financial markets, particularly during COVID [@pagano2020]. Users of Robinhood increased the amount of money they invested on the platform during the pandemic [@welch2020]. Robinhood traders primarily engage in both momentum and contrarian strategies where they invest in stocks that have already demonstrated rising price momentum and "buying the dip" where they buy stocks that have recently fallen in price, respectively [@pagano2020]. In other words, Robinhood investors did not panic during broad market turmoil during COVID-19 and often used price drops as buying opportunities. However, in aggregate they do not produce an alpha, or "beat the market" by producing larger-than-average returns [@eaton2021]. Their trades ultimately mostly produce noise in financial markets [@eaton2021]. If all Robinhood investors do is create "noise" in the markets, then the noisiest they've been is when GameStop stock rose dramatically in early 2021. 

## GameStop

In January and February 2021, a handful of stocks experienced an extreme "short-squeeze" which effectively sky-rocketed the prices of these stocks that was caused by decentralized retail investors acting in concert [@lyocsa2021]. The largest movement was seen in GameStop which took most of the attention of the saga. Many investors, both individual and institutional, gained and lost significant sums of money. For a full recapping of the sequence of events, I recommend _________. 

The GameStop saga provides support for the principle that market forces tend to make markets fair, "where fairness is defined as investors 'getting what they pay for' rather than investors 'beating the market'" due to the apparent misvaluation of GameStop and other stocks by institutional investors [@macey2021]. On the other hand, the events also cast doubt as to whether the SEC is achieving its stated goal of maintaining fair, orderly, and efficient markets" [@macey2021].

The retail investors who caused the short squeeze organized on a subforum of the popular website of Reddit called "Wall Street Bets," or r/wallstreetbets [@lyocsa2021]. As a result, the GameStop saga is seen as battle between Wall Street and small, Robinhood traders who are upset at Wall Street. The members of r/wallstreetbets maintained a sense of unity and purpose throughout the events [@dimuzio2021]. Regardless of whether the organized behavior of r/wallstreetbets users is a political movement, hopeful acts of "sticking it to hedgefunds" or something else, they have demonstrated their power to move financial markets [@dimuzio2021]. However, retail investors as whole both bought and "shorted"--or bet that GameStop stock was going to go down--indicating that "the Gamestop frenzy was not a pure digital protest against Wall Street but speculative trading by a group of retail investors, in line with their prior high-risk trading behavior" [@hasso2021].

## r/WallStreetBets 

Wall Street Bets is a subreddit where investors talk about their stock trades. The group is known for their risky "YOLO" (You Only Live Once) trades [@dimuzio2021]. There is a particular language and meme culture that permeates from the group and potentially helps to create their culture of "degeneracy" [@boylston2021]. For example, they celebrate people with "diamond hands" as those who are prepared to hold to their stocks for a long time and decry "paper hands," those who miss out on profits by selling too a stock too early [@wsbshop]. They maintain they are not "dumb," but they are "retarded" or "autists." Although these words are derogatory, they use it as a way to self-deprecate and build unity.

# Information and Media

Scholars have been particularly interested with the role of information on financial markets. Business researchers largely operate under the efficient market hypothesis where markets reflect accurate valuations based on available information. As such, they are primarily interested in whether information being spread about companies is timely or "stale" and the effects of media attention on stocks, media including television, newspaper, digital news, and social media on financial markets. Overall, media attention-driven behavior favors individuals buying stocks opposed to selling stocks [@barber2007].

## Newspapers

Even though one may think of the stock market reacting to news instantaneously, media attention given to individual stocks via newspapers impact the pricing of the stock. The causal link between newspapers impacting trading prices and volumes was established by leveraging newspaper strikes and observing the subsequent impact on financial markets [@peress2014]. This effect persists for many days after public, and the impact is prolonged during a recession compared to an expansion [@antweiler2006]. However, not all coverage is good coverage. Stocks with little or no attention in newspapers earn higher returns than high-media-attention stocks [@fang2009].

Stock market investors respond differently to new versus stale news--as defined by textual similarity of sequential newspaper stories. The prices of stocks respond less to stale news, but there is a reversal in the movement of a stock's price after stale news, suggesting that initial movements are overreactions [@tetlock2011]. Further, quantitative information is easier for news readers to process compared to qualitative information [@engelberg2008].

In addition, it is not just substantive news about a stock that impacts performance, stock recommendations in the _Wall Street Journal's_ "Dartboard" column predicted abnormal positive returns of 4% and double the trading volume over the two days after the column recommended a stock [@barber1993]. These market abnormalities are largely driven by naive investors and are ultimately noise in the financial markets. Price movements from the column were reversed within 15 days, and investors following the "expert" recommendations of the column lost 3.8% on a risk-adjusted basis in the six months following the recommendation [@liang1999]. These findings show that media content has a clear relationship with asset prices and should not be seen as a "sideshow," are consistent with the theoretical models of retail investors acting as noise in the financial ecosystem, and is inconsistent with the theory that media can serve as a proxy for new financial information that informs the efficient market on fundamental asset valuations [@tetlock2007]. 

## Television

Similar to newspapers, the financial markets react to media content on television, often in real-time. When the TV program "The Morning Call" on _CNBC_ reports analyst views on individuals stocks, the market reacts within seconds, the price is fully incorporated within one minute, and trading intensity in the stock doubles within that first minute [@busse2002]. "Mad Money with Jim Cramer" is a popular _CNBC_ television show where the host, Jim Cramer, makes a bevy of buy and sell recommendations on stocks. There are significant price movements for stocks that Cramer recommends to buy and sell [@bolster2012; @engelberg2012; @karniouchina2009]. The effects on the movement of stocks given a "buy" rating from Cramer are quickly reversed, but the effect persists longer for "sell" recommendations [@bolster2012]. The viewing ratings of the show even predicts the strength of the price movements [@engelberg2012]. Further, traditional advertising variables, such as message length, recency-primacy effects, information clutter, and source credibility all influence the size of the market reaction to a "buy" recommendation from Cramer [@karniouchina2009], which potentially suggests that content consumers process media from financial pundits similarly to other types of advertising. Ultimately, these effects show that the investors taking actions from Cramer's recommendations are uninformed traders and not experienced investors receiving any new information [@keasler2010].

## Social Media

As evidenced by the recent GameStop and r/wallstreetbets saga, social media is playing an increasing role in the way that individual investors collect, share, and process information that results in investment decisions. Literature on social media and the stock market fall into three broad categories: social media as a predictor of stock movements, crowd-sourced financial recommendation, and behavioral contagion/ herding. 

First, social media has been used to try to predict fluctuations in individual stocks--with mixed results. The number of posts on TheRagingBull.com, one of the original stock message boards, in 2001 coincided with days with abnormally high trading volume, but they were not predictive of market pricing or volume after for controlling for industry-adjusted returns or normal trading volume [@tumarkin2001]. This result suggests that some other exogenous event is a confounder that drives both social media engagement and stock market volume. However, individual tweets just before a firm's quarterly earnings announcement is a predictor of the earnings report and subsequent price action [@bartov2018]. News diffusion on Twitter leads to lower bid-ask spread and price pressure during a news day, but the effect is reversed the next day [@chawla2016]. Twitter, like traditional media, spreads stale news, albeit at a higher speed than traditional media [@chawla2016]. 

Other new media, including Yahoo! stock message boards [@das2007] and Google Search Volume [@da2011] are all predictive of stock movements. Even disagreements on StockTwits, a Twitter-like stock-based social media, predicts price changes by affecting trading done by uninformed investors and facilitates trading by informed traders who are taking actions aimed at changing corporate policies [@cookson2021]. 

Next, social media have been studied as way to crowd source financial advice. For example, Seeking Alpha is a website where non-professional investors can write articles analyzing a stock. Seeking Alpha articles are predictors of stock market movements [@chen2014; @farrell2018]. Estimize is an open platform that solicits forecasts from contributors. These crowdsourced forecasts are incrementally useful in forecasting earnings, and the higher the volume of forecasts, the more accurate the predictions are [@jame2016]. SumZero.com is a private social networking site that facilitates interaction and information exchange among professional investors. Recommendations offered on this website generate significant returns in the financial markets, particularly, long-term returns are particularly strong for "contrarian" buy recommendations [@crawford2018]. 

# References